Expert-In-The-Loop Causal Discovery: Iterative Model Refinement Using Expert Knowledge

### 2025-02-08 paper/main.tex
Adds the ranked expand algorithm
Updates

\subsection{Ranking Potential New Edges}

@@

As the automated algorithms can only recover the CPDAG from data, and our
method recovers the DAG. To compare them fairly, we computed the SHD and SID
for all possible orientations of the CPDAG to show the best and worst case
scenarios. The results are shown in Figure~\ref{shd_sid}. For SHD, the
performance of \emph{ExpertInLoop} is comparable to the automated algorithms
for $ \alpha = 0.3 (\alpha_{\textrm{eff}} = 0.53 $, and performs better for
higher $ \alpha $ values. For SID, the \emph{ExpertInLoop} performs better in
denser DAGs.

>>>>

In this section, we give some of the practical aspects of the
\emph{ExpertInLoop} algorithm, and then compare the algorithm with automated
causal discovery algorithms.

\subsection{Ranking Potential New Edges}
In Algorithm~\ref{algo:expert}, in each step we add a single edge to the DAG
and remove any edges that can be pruned. In practice, we use a slightly
modified version of the Algorithm~\ref{algo:expand} for adding new edges as
given in Algorithm~\ref{algo:ranked_expand}. Rather than choosing the first
potential edge addition, this algorithm ranks all potential new edges based on
the conditional association between the variables. The edge with the highest
conditional association is then added to the graph. This modification allows us
to prioritize edges that would help in the explaining the observed correlation
in the data the highest. For this purpose, we use a measure of conditional
association measure, $\phi$ between variables.

\begin{algorithm}[h]
\DontPrintSemicolon
\SetAlgoLined
\SetKwFunction{RankedExpand}{RankedExpand}
\SetKw{KwGoTo}{go to}
\SetKwProg{Fn}{Function}{:}{}
\Fn{{\sc RankedExpand}($V,E,\mathcal{D}_G,\mathcal{A}_G$,$\phi$,$B=\emptyset$)}{
    let $\phi_{\max} \gets 0 $\;
    \ForEach{$X, Y$ where $X \to Y \notin E$ and $Y \to X \notin E$ and 
	$\{X,Y\} \notin B$}{
        let $\mathbf{Z}$ be a set that $d$-separates $X$ and $Y$ in $(V,E)$\;
        \If{$\mathcal{D}_G(X, Y, \mathbf{Z}) = 0$}{
		\If{$\lvert \phi(X, Y, \mathbf{Z}) \rvert > \phi_{\max}$}{
			$ \phi_{\max} \gets \lvert \phi(X, Y, \mathbf{Z}) \rvert $ \;
			$ L_{\max} \gets \mathcal{A}_G(X, Y)$ \;
		}
        }
    }
    add $L_{\max}$ to $E$\;
    \Return{$(V,E)$}\;
}
\caption{Adding an edge between variables with the highest correlation}
\label{algo:ranked_expand}
\end{algorithm}

@@

As the automated algorithms can only recover the CPDAG from data, and our
method recovers the DAG. To compare them fairly, we computed the SHD and SID
for all possible orientations of the CPDAG to show the best and worst case
scenarios. The results are shown in Figure~\ref{fig:shd_sid}. For SHD, the
performance of \emph{ExpertInLoop} is comparable to the automated algorithms
for $ \alpha = 0.3 (\alpha_{\textrm{eff}} = 0.53 $, and performs better for
higher $ \alpha $ values. For SID, the \emph{ExpertInLoop} performs better in
denser DAGs.
